Applying Radial Basis Networks and Markov Chains for on-line detection of concept drift in non-stationary environments

### 2019-10-27 tex/article/article.tex
Complement abstract

\begin{abstract}
Most real-world problems experience a phenomenon known as concept drift,
which is a change in data distribution that can affect the system performance.
%
However, the majority of drift detection methods are unfit for non-stationary environments with data streams.
These algorithms or require the correct labeling of data and detect drift through the monitoring of the system performance,
or do not attend the severe response time and resource usage restrictions inherent to data streams.
%
In an attempt to mitigate the aforementioned problem,
this paper proposes a novel proactive method for on-line drift detection,
called RBFChain.
The proposed method relies on Radial Basis Networks implicit clustering property,
besides using Markov Chains to model the transitions and achieve noise tolerance.
%
To assess the proposed method as a viable concept drift detector,
an analysis of sensitivity, accuracy,
and noise tolerance was performed using synthetic datasets,
and results were compared to the most established algorithms in the literature.
%
Furthermore, the algorithm was applied to the real-world problem of eye-tracking.
A relevant problem with impact in different areas of knowledge, since many behavioral experiments use eye-tracking information (fixations and saccades) as a relevant analysis factor.
Performed experiments reveal that RBFChain can classify fixations and saccades in real-time, with high accuracy, noise tolerance, and using limited resources.
\end{abstract}

>>>>

\begin{abstract}
Most real-world problems experience a phenomenon known as concept drift,
which is a change in data distribution that can affect the system performance.
%
However, the majority of drift detection methods are unfit for non-stationary environments with data streams.
These algorithms or require the correct labeling of data (unviable in these settings) and detect drift through the monitoring of the system performance,
or do not attend the severe response time and resource usage restrictions inherent to data streams.
%
In an attempt to mitigate the aforementioned problem,
this paper proposes a novel proactive method for on-line drift detection,
called RBFChain.
The proposed method relies on Radial Basis Networks implicit clustering property,
besides using Markov Chains to model the transitions and achieve noise tolerance.
%
To assess the proposed method as a viable concept drift detector,
an analysis of sensitivity, accuracy,
and noise tolerance was performed using synthetic datasets,
and results were compared to the most established algorithms in the literature.
%
Furthermore, the algorithm was applied to the real-world problem of eye-tracking.
A relevant problem with impact in different areas of knowledge, since many behavioral experiments use eye-tracking information (fixations and saccades) as a relevant analysis factor.
Performed experiments reveal that RBFChain can classify fixations and saccades in real-time, with high accuracy, noise tolerance, and using limited resources.
\end{abstract}
